Using spatially biased variables in ecosystem condition accounting with a GIS based workflow

### 2024-02-22 manuscript/HIAs.qmd
geom_taxt inside bars

```{r}
#| eval: false
# The year effect is constant, so ignoring that
corrCheck |>
  mutate(
    i_ADSV_fct = floor(round(i_ADSV * 10, 2)) / 10,
    i_alien_fct = floor(round(i_alien * 10, 2)) / 10,
    i_ditch_fct = floor(round(i_ditch * 10, 2)) / 10
  ) |>
  pivot_longer(
    cols = c(i_ADSV_fct, i_alien_fct, i_ditch_fct),
    values_to = "indicatorValue",
    names_to = "indicator",
    values_drop_na = T
  ) |>
  ggplot(aes(
    x = factor(infrastructureIndex),
    fill = factor(indicatorValue)
  )) +
  geom_bar(
    position = "fill"
  ) +
  theme_bw(base_size = 12) +
  guides(fill = guide_legend("Scaled indicator values")) +
  ylab("Fraction of data points") +
  xlab("HIA") +
  scale_fill_brewer(
    palette = "RdYlGn") +
  facet_grid(~indicator)
```

```{r}
#| eval: true
# Using a color gradient emphasizes the first color (dark green).
# Lets try discrete colors, and merge som classes to simplify
(validationPlot <- corrCheck |>
  #rename(ADSV = i_ADSV) |>
  pivot_longer(
    cols = c(i_ADSV, i_alien, i_ditch),
    values_to = "indicatorValue",
    names_to = "indicator",
    values_drop_na = T
  ) |>
  mutate(
    condition = case_when(
      indicatorValue < 0.6 ~ "<0.6",
      indicatorValue < 0.8 ~ "0.6 to 0.8",
      indicatorValue < 0.91 ~ "0.8 to 0.9",
      .default = "0.9 to 1"
    ),
    condition = fct_reorder(condition, indicatorValue),
    indicator = case_when(
      indicator == "i_ADSV" ~ "ADSV",
      indicator == "i_alien" ~ "Alien species",
      indicator == "i_ditch" ~ "Trenching"
    )
  ) |>
  ggplot(aes(
    x = factor(infrastructureIndex),
    fill = condition
  )) +
  geom_bar(
    position = "fill"
  ) +
  theme_bw(base_size = 12) +
  guides(fill = guide_legend("Indicator values")) +
  ylab("Fraction of data points") +
  xlab("Human Impact Areas") +
  scale_fill_brewer(palette = "RdYlGn") +
  facet_grid(~indicator)
)
```

@@

(methodsMap <- tmap_arrange(
  positionMap,
  muniPlot,
  infraMuniMap_l,
  infraMuniMap,
  widths = c(.35, .65),
  heights = c(.6, .4),
  outer.margins = NULL
))

>>>>

```{r}
#| eval: false
# The year effect is constant, so ignoring that
corrCheck |>
  mutate(
    i_ADSV_fct = floor(round(i_ADSV * 10, 2)) / 10,
    i_alien_fct = floor(round(i_alien * 10, 2)) / 10,
    i_ditch_fct = floor(round(i_ditch * 10, 2)) / 10
  ) |>
  pivot_longer(
    cols = c(i_ADSV_fct, i_alien_fct, i_ditch_fct),
    values_to = "indicatorValue",
    names_to = "indicator",
    values_drop_na = T
  ) |>
  ggplot(aes(
    x = factor(infrastructureIndex),
    fill = factor(indicatorValue)
  )) +
  geom_bar(
    position = "fill"
  ) +
  theme_bw(base_size = 12) +
  guides(fill = guide_legend("Scaled indicator values")) +
  ylab("Fraction of data points") +
  xlab("HIA") +
  scale_fill_brewer(
    palette = "RdYlGn") +
  facet_grid(~indicator)
```
```{r}
temp <- corrCheck |>
  pivot_longer(
    cols = c(i_ADSV, i_alien, i_ditch),
    values_to = "indicatorValue",
    names_to = "indicator",
    values_drop_na = T
  ) |>
  mutate(
    condition = case_when(
      indicatorValue < 0.6 ~ "<0.6",
      indicatorValue < 0.8 ~ "0.6 to 0.8",
      indicatorValue < 0.91 ~ "0.8 to 0.9",
      .default = "0.9 to 1"
    ),
    condition = fct_reorder(condition, indicatorValue),
    indicator = case_when(
      indicator == "i_ADSV" ~ "ADSV",
      indicator == "i_alien" ~ "Alien species",
      indicator == "i_ditch" ~ "Trenching"
    )
  ) |>
  as_tibble() |>
  group_by(indicator, infrastructureIndex, condition) |>
  summarise(n = n()) |>
  ungroup() |>
  group_by(indicator, infrastructureIndex) |>
  mutate(lab = round(n/sum(n)*100),
         lab = case_when(
           lab < 5 ~ NA,
           .default = paste0(lab, "%")
         ))

nrow(temp)

temp |>
ggplot(aes(
    x = infrastructureIndex,
    y = n,
    fill = condition
  )) +
  geom_bar(
    position = "fill",
    stat = "identity"
  ) +
  geom_text(aes(label = lab),
    position = position_fill(vjust = 0.5),
    color= "black", vjust = 0.5, size = 4) +
  theme_minimal(base_size = 15) +
  theme(
    panel.grid = element_blank(), 
    axis.text.x = element_text(margin = margin(t = -10)),
    axis.text.y = element_blank(),
    strip.text = element_textbox(
      size = 12,
      halign = 0.5, linetype = 1, r = unit(5, "pt"), width = unit(1, "npc"),
      padding = margin(2, 0, 1, 0), margin = margin(3, 3, 3, 3))) +
  guides(fill = guide_legend("Indicator values")) +
  ylab("Fraction of data points") +
  xlab("Human Impact Areas") +
  scale_fill_manual(values = c("#E85437","#FBBF00", "#B5DF73", "#009000")) +
  facet_grid(~indicator)

```{r}
#| eval: true
# Using a color gradient emphasizes the first color (dark green).
# Lets try discrete colors, and merge some classes to simplify
(validationPlot <- corrCheck |>
  #rename(ADSV = i_ADSV) |>
  pivot_longer(
    cols = c(i_ADSV, i_alien, i_ditch),
    values_to = "indicatorValue",
    names_to = "indicator",
    values_drop_na = T
  ) |>
  mutate(
    condition = case_when(
      indicatorValue < 0.6 ~ "<0.6",
      indicatorValue < 0.8 ~ "0.6 to 0.8",
      indicatorValue < 0.91 ~ "0.8 to 0.9",
      .default = "0.9 to 1"
    ),
    condition = fct_reorder(condition, indicatorValue),
    indicator = case_when(
      indicator == "i_ADSV" ~ "ADSV",
      indicator == "i_alien" ~ "Alien species",
      indicator == "i_ditch" ~ "Trenching"
    )
  ) |>
  ggplot(aes(
    x = factor(infrastructureIndex),
    fill = condition
  )) +
  geom_bar(
    position = "fill"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    panel.grid = element_blank(), 
    axis.text.x = element_markdown(),
    axis.text.y = element_blank()) +
  guides(fill = guide_legend("Indicator values")) +
  ylab("Fraction of data points") +
  xlab("Human Impact Areas") +
  scale_fill_brewer(palette = "RdYlGn") +
  facet_grid(~indicator)
)
```

@@

(methodsMap <- tmap_arrange(
  positionMap,
  muniPlot,
  infraMuniMap_l,
  infraMuniMap,
  widths = c(.35, .65),
  heights = c(.6, .4),
  outer.margins = .01
))
